Collaborative analysis and writing in R

### 2017-12-31 TUTORIAL.Rmd
Finished tutorial

---
title: "Tutorial"
author: "R Clay Wright"
date: "12/28/2017"
output: html_document
---

@@

This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled A comprehensive analysis of natural sequence variation within the <i>Arabidopsis thaliana</i> nuclear auxin response pathway" (running title: "nuclear auxin response natural variation""). I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio, 
but if you feel you could use a refresher check out [this introduction](https://github.com/wrightrc/Intro_to_R_RStudio_and_ANOVA). We'll use docker to share with you a version of R and Rstudio that already has all of the necessary packages to do the analysis. We'll use Rmarkdown to write the manuscript and make figures and tables by integrating prose, R code and results from the code.

So in this tutorial I will guide you through in as abstracted a way as possible:
1. Installing docker
2. Starting the version of RStudio in docker 
3. Working in docker/RStudio
4. Saving your work in docker/RStudio
5. Using Rmarkdown to format text
6. Using Rmarkdown to insert R code and results into text
7. Citations in Rmarkdown

@@

If you click on the docker icon in the menu bar you will see kitematic listed as an option.  
![](TUTORIAL_files/Screenshot 2017-12-29 16.01.58.png)  
Select "Kitematic" to open a window with a link to download and install kitematic. Once you have installed kitematic selecting the "Kitematic" link will open the kitematic interface.  
![](TUTORIAL_files/Screenshot 2017-12-29 16.50.28.png)  
The containers that you currently have running on your system are listed in the left panel. You can see I have a "nuclear-auxin-natural-variation" container running, that docker has given the creative name "distracted-nobel". In the main pane of kitematic there are several popular images that we can create new containers from.

@@

You can search in the box at the top of kitematic for "r1001genomes", then click "create" to download the image and create a container. This will take a **while**. You are downloading linux, R, Rstudio and all of those packages. Fortunately docker will reuse all of these components whenever you download a related image.  
![](TUTORIAL_files/Screenshot 2017-12-29 21.08.01.png)

Once the image is downloaded and the container is created you'll see a log and some other information "IP & PORTS" and "VOLUMES".  
![](TUTORIAL_files/Screenshot 2017-12-29 21.13.24.png)

@@

Go ahead give it a try!  
![](TUTORIAL_files/Screenshot 2017-12-29 21.34.22.png)  
Now we need some data to play with in our Rstudio container! You'll notice that the only folder available in the "Files" pane is "kitematic", which matches the folder in the "VOLUMES" pane back in kitematic. To make use of this folder we need to map it to a folder on our computer. Click on the "gear" settings icon in the "VOLUMES" pane, then click "CHANGE" and select a folder to be able to access it within your container. Go ahead and select the "natural-variation-webtool" folder that this tutorial is in.

@@

This document was is actually written in R Markdown. If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane.  
![](TUTORIAL_files/Screenshot 2017-12-29 22.45.46.png)

@@

## Including Plots

@@

I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)! You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu.  
![](TUTORIAL_files/Screenshot 2017-12-29 22.57.15.png)

>>>>

---
title: "Collaborative analysis and writing in R"
author: "R Clay Wright"
date: "12/28/2017"
output: html_document
---

@@

<!--This is a CSS (Cascading Style Sheet).
It controls how panels and images are formatted.></!-->
<style>
  img {
    display: block;
    margin: auto;
    width: 90%;
    height: auto;

  }
</style>

This tutorial will (hopefully) gently guide you through how to use the tools docker and Rmarkdown to analyze the natural variation in your set of genes and collaborate on our manuscript, tentatively titled: *A comprehensive analysis of natural sequence variation within the Arabidopsis thaliana nuclear auxin response pathway* (running title: " natural variation in the nuclear auxin response"). I assume that you have a basic working knowledge of how to download and install software as well as R and RStudio, 
but if you feel you could use a refresher check out [this introduction](https://github.com/wrightrc/Intro_to_R_RStudio_and_ANOVA). We'll use docker to share with you a version of R and Rstudio that already has all of the necessary packages to do the analysis. We'll use Rmarkdown to write the manuscript and make figures and tables by integrating prose, R code and results from the code.

So in this tutorial I will guide you through, in as abstracted a way as possible:  
1. Installing docker  
2. Starting the version of RStudio in docker  
3. Working in docker/RStudio  
4. Saving your work in docker/RStudio  
5. Using Rmarkdown to format text  
6. Using Rmarkdown to insert R code and results into text  
7. Citations in Rmarkdown

@@

If you click on the docker icon in the menu bar you will see kitematic listed as an option.  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 16.01.58.png)  
Select "Kitematic" to open a window with a link to download and install kitematic. Once you have installed kitematic selecting the "Kitematic" link will open the kitematic interface.  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 16.50.28.png)  
The containers that you currently have running on your system are listed in the left panel. You can see I have a "nuclear-auxin-natural-variation" container running, that docker has given the creative name "distracted-nobel". In the main pane of kitematic there are several popular images that we can create new containers from.

@@

You can search in the box at the top of kitematic for "r1001genomes", then click "create" to download the image and create a container. This will take a **while**. You are downloading linux, R, Rstudio and all of those packages. Fortunately docker will reuse all of these components whenever you download a related image.  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.08.01.png)

Once the image is downloaded and the container is created you'll see a log and some other information "IP & PORTS" and "VOLUMES".  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.13.24.png)

@@

Go ahead give it a try!  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 21.34.22.png)  
Now we need some data to play with in our Rstudio container! You'll notice that the only folder available in the "Files" pane is "kitematic", which matches the folder in the "VOLUMES" pane back in kitematic. To make use of this folder we need to map it to a folder on our computer. Click on the "gear" settings icon in the "VOLUMES" pane, then click "CHANGE" and select a folder to be able to access it within your container. Go ahead and select the "natural-variation-webtool" folder that this tutorial is in.

@@

This document was is actually written in R Markdown. If you haven't already, open the "code" markdown version of this document ("TUTORIAL.Rmd") in Rstudio, and navigate down to the "R Markdown" section using the navigation bar at the bottom of the "Source" pane.  
![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.45.46.png)

@@

### Including Plots

@@

Don't get bogged down in formatting the text, but I would highly recommend checking out the "more details" link above and don't miss the [cheatsheet](http://rmarkdown.rstudio.com/lesson-15.html)! You can also find this and other cheatsheets and a Markdown reference in the RStudio "Help" menu. Focus foremost should be on the content. We'll worry about formatting in revisions.
![](TUTORIAL_screenshots/Screenshot 2017-12-29 22.57.15.png)

In the manuscript skeleton and the example analysis you can also see some examples of how to use R markdown, DT (data table) and kable to format and insert tables into our document.

### Citations in R markdown  
You can include citations in markdown documents using the following syntax: `[@citation_key]`. The citation key refers to the unique identifier generated when exporting a bibliography from a reference manager. For the manuscript you will need to export the references from your reference editor in one of many formats, but I would recommend BibTex (.bib). For other formats see [the markdown documentation](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#citations).

Let's take a look at an example, open the "references.bib" file in the "manuscript folder". You'll notice each reference begins with an @-symbol and the type of reference. Then within curly braces are all of the data about the reference. The first item is the citation key, which in this case is "gray_identification_1999". So to cite this paper in markdown you would add `[@gray_identification_1999]` to the end of a sentence. My process for citing-while-writing is to first write, then when I want to cite something, find the paper in my reference manager and copy it to a folder for the project I'm working on. This will allow me to export the folder as a .bib file later). Then I copy the citation key for the reference and paste it into my writing. When I'm ready to knit the document, I'll export the folder of references as a .bib (with Better Bibtex you can sync this file so that it is automatically updated). This bibliography file must be in the same folder with the .Rmd file, along with a .csl file that specifies the format of the citations and bibliography in the final document (available for most journals [via Zotero](https://www.zotero.org/styles)). These files also must be specified in the header of the document.  
![](TUTORIAL_screenshots/Screenshot 2017-12-30 14.04.14.png)

Most common reference managers allow you to export a bibliography as a bibtex file, and copy and paste citation keys, however I would highly recommend downloading [Zotero](www.zotero.org) and importing your reference library from whatever reference editor you currently use. Zotero is free, open-source and integrates very well with all popular web browsers and writing tools. If you want to work in a different citation manager, here are some links to help you create and copy and paste citation keys as well as export BibTex bibliography files. Note that most of these tutorials are for Latex, which uses a slightly different syntax from Rmarkdown. In Rmarkdown insert citations via `[@citation_key]`. The `citation_key` string does have some stipulations 
>The citation key must begin with a letter, digit, or \_, and may contain alphanumerics, _, and internal punctuation characters (:.#$%&-+?<>~/).

*To insert multiple citations separate the two citations by semicolons, e.g.* `Blah blah [@smith04; @doe99].`

[Endnote](https://www.reed.edu/cis/help/LaTeX/EndNote.html): Export as [bibtex with labels](http://www.endnote.com/downloads/style/bibtex-export-using-en-label-field).  
Endnote seems to have some issues generating unique citation keys for each reference. It might be best to export your library as an Endnote XML and import the whole thing into Zotero, if you have issues with the above solutions.  
[Mendeley](https://blog.mendeley.com/2011/10/25/howto-use-mendeley-to-create-citations-using-latex-and-bibtex/)  
[Zotero](https://github.com/bbest/rmarkdown-example#writing-with-rmarkdown)

Once you are done writing and citing and have exported your BibTex file, you can add your bibliography file to the manuscript folder and add the name of your file to the document header (comma separated). This should allow your version of the manuscript to knit, with perhaps some errors if there are duplicate citation keys. Don't worry we'll sort out duplicates later, once I have merged everyones sections' together.

#### Also print out your key citations and we will make a binder for everyones use in the computer room! Also don't forget to send me any changes or suggestions you have for this tutorial.

Thanks and happy collaborative writing!
